---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# PCAssess

<!-- badges: start -->
<!-- badges: end -->

PCAssess is an R package for assessing high grading bias in population genetics. PCAssess implements and automates permuation tests to detect high grading bias in Principal Component Analysis (PCAs). 

## Installation

You can install the development version of PCAssess from [GitHub](https://github.com/hemstrow/PCAssess) with:

``` r
# install.packages("pak")
pak::pak("hemstrow/PCAssess")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(PCAssess)
## basic example code
#perm_res <- run_permutation(mon_sn, "pop", 10, fst_cut=.95, par = FALSE, store_pca = FALSE) # run permutation test on example Monarch data
#plot_permutation_res(perm_res, 10) # plot observed PCA and 100 permutations
#plot_permutation_res(perm_res, 0) # plot only the observed PCA
```
