---
title: "Asessing highgrading bias in genetic data with PCAssess"
author: "Andy Lee and William Hemstrom" 
date: "Feburary 10, 2025" 
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(PCAssess)
set.seed(1213)
```

# PCAssess Introcution
PCAssess is an R package for assessing high grading bias in population genetics. PCAssess implements and automates permuation tests to detect high grading bias in Principal Component Analysis (PCAs). 

# Quick-start Guide
Here, we will walk through a basic analysis of the example monarch butterfly data included in this package (mon_sn). We will run permutation tests and visualize the results. 

# Data import
PCAssess requires two inputs. First, genetic data formatted as single numbers indicated genotype, where 0 and 2 are homozygoytes and 1 is a heterozygote. Rows are SNPs and columns are individuals. Second, a character vector containing categorical variable by which the data was broken up e.g., population, treatment.
```{r}
snps_dat <- system.file("extdata", "mon_sn.rda", pcakage = "PCAssess")
```

# Running permutation tests
This example runs permutation test on mon_sn, the example dataset included in the package. In this dataset, samples were separated into various populations "pop". Populations code are permuted 100 times, and top 5% FST loci are chosen with the default setting, 0.95. The resulting object, res, is an is a list containing: null distribution of the F-statistic observed values, the observed F-statistic, and the p-values from PCA permutation testing
```{r}
# 
res <- run_permutation(mon_sn, "pop", 100, fst_cut = , par = FALSE, store_pca = FALSE) 
```

# Visualize permutation results
This example visualizes the results from the above permutation tests. 
```{r}
plot_permutation_res(perm_res, 100) # plot distribution of F-statistic from 100 permutations
plot_permutation_res(perm_res, 0) # plot only the observed PCA
```

